Carbon stocks in aboveground biomass for Colombian mangroves with associated uncertainties

### 2018-01-06 LaTeX_source_files/Article/Article_mangroves.tex
figures were ereased from latex in folder for journal

\begin{abstract}
Decision-making process as well as the implementation of future climate change mitigation strategies for mangrove ecosystems, need the availability of reliable base information and tools to estimate the carbon stocks at regional and national levels. We estimated carbon (C) stocks in aboveground biomass (AGB) for mangroves in the Caribbean and Pacific coasts of Colombia. Using available data on AGB density and mangrove area for the whole country (excluding islands) and each coast independently, we estimated a national carbon stock in AGB for Colombian mangroves as 14.95 $\pm$ 2.72 TgC (mean $\pm$ SE), with 2.20 $\pm$ 0.86 TgC in the Caribbean coast and 9.61 $\pm$ 2.78 TgC in the Pacific coast. Uncertainty for total carbon in AGB in Colombian mangroves, reported as SE/mean in percentage, was 18\% at the national level, 39\% in the Caribbean coast, and 29\% in the Pacific coast. This uncertainty was more influenced by uncertainties associated with the estimation of mangrove area for the Caribbean coast, while for the Pacific coast it was more influenced by the uncertainties associated with AGB density. This difference is the result of a contrasting availability of AGB density data for both coasts.  Comparison between observed AGB density data and predictions from large-scale models showed that these models underestimate AGB density for Colombian mangroves. We reparameterized these models with our data, but found poor goodness-of-fit statistics for these model structures. We propose therefore three new statistical models to predict AGB density in Colombian mangroves based on climatic and vegetation data. In all cases, the best models included the enhanced vegetation index (EVI) and the mean temperature of driest quarter (BIO9). 
\end{abstract}

@@

The final number of sources used reflects the low availability of public information in Colombia regarding biomass studies in mangrove ecosystems. Only those studies published with open access, or referenced in the Information System for the Management of Mangroves in Colombia (SIGMA) were used. Additionally, despite more biomass inventories have been conducted in the Colombian mangroves, only those provided by INVEMAR could be accessed. Given the differences in terms of scale of the two sources of information, as well as the impossibility to encompass the different mangrove types along the two coasts, there is a limitation in terms of the scope of our results. Nevertheless, considering the goal of the study, this first approach is a relevant step to identify the neccesities of sampling in terms to reduce the current uncertainties associated to the carbon stock estimation at national and/or regional level.

@@

%\begin{figure}[htbp]
%   \centering
%   \includegraphics{Figures/map.pdf} % requires the graphicx package
%   \caption{Spatial distribution of mangroves in Colombia (green dots) and location of study sites with AGB density information (red stars).}
%   \label{fig:map}
%\end{figure}
%
Furthermore, information about mangrove cover area comes from two different sources: global estimations \citep{FAO2007, Giri2011, Giri2011dataset, CONL:CONL12060} and national estimations compiled by the Colombian institute for ocean and coastal research \citep{INVEMAR2014} and derived from particular estimations made by the different environmental authorities with jurisdiction in mangrove areas along the country. Those mangrove areas were estimated under different methodological approaches.

@@

Total carbon in AGB showed larger values in Pacific mangroves (9.61 $\pm$ 2.78 TgC) (mean $\pm$ SE) than in Caribbean mangroves (2.20 $\pm$ 0.86 TgC). This result is highly influenced by mangrove area, because if the comparison is made in terms of the same area unit (hectare), values of carbon density between both coasts do not show significant differences (Figure \ref{fig:Csotck}).

\begin{figure}
\includegraphics[%scale=0.4
width=\textwidth
]{Figures/Cdensity_and_TotalCstock.pdf} 
\caption{
Aboveground biomass (AGB) density (Mg C ha$^{-1}$) and total carbon stock in AGB for Colombian mangroves (Tg C). 
}
\label{fig:Csotck}
\end{figure}

@@

To achieve a 10\% uncertainty in AGB density for the Pacific coast, it would be necessary to establish at least 47 new sampling units, and a 5\% uncertainty may not be possible to achieve with less than 200 sampling units (Figure \ref{fig:samplingEffort}). For the Caribbean coast the situation is less critical; it is possible to achieve a 5\% uncertainty in AGB density by adding 73 additional sampling units. Combined for the entire country, 81 additional sampling units are required to achieve 5\% uncertainty in AGB density. In all cases, we assume that these additional sampling units must follow the principles of random sampling.

@@

\begin{figure} 
\includegraphics[%scale=0.4
width=\textwidth
]{Figures/AGBsamplingEffort.pdf} 
\caption{
Number of necessary sampling units of AGB density required to reduce current uncertainties at 10\% and 5\% level for Caribbean, Pacific and total Colombian mangroves.} \label{fig:samplingEffort}
\end{figure}

@@

We tested the ability of the models proposed by \citet{Twilley1992} and \citet{CONL:CONL12060}  in predicting our AGB density data. For both models, we found important overestimations of AGB density for the majority of the observations, although in some cases the models underestimated observed AGB density, particularly for values larger than 150-200 Mg ha$^{-1}$  (Figure \ref{fig:globalModels}).

@@

\begin{figure} 
\includegraphics[%scale=0.4
width=\textwidth
]{Figures/Globalmodelsvs.pdf} 
\caption{
Observed vs. predicted AGB density (Mg ha$^{-1}$). Predicted values: using global models \citep{Twilley1992, CONL:CONL12060} and observed data: current data available for mangrove areas in Colombia} \label{fig:globalModels}
\end{figure}

@@

\begin{figure} 
\includegraphics[%scale=0.4
width=\textwidth
]{Figures/AGBdensity_map_final.pdf} 
\caption{
Map of predicted AGB density (Mg ha$^{-1}$) based on the candidate models classified as ``substantial weight models''} \label{fig:Predicted AGB density map}
\end{figure}

@@

%In general terms, to reduce uncertainties related with total AGB carbon stock estimation at coast level as well as at national level we propose the following ways in terms of area and AGB efforts, according with \citet{IPCC2003} recommendations:
%
Regarding to area estimation, the use the remote sensing with the ground based surveys approach \citep{IPCC2003} has the advantage to be  the most cost cost-efficient method of area measurement. Furthermore, to reduce current uncertainties it is important to check for consistent relationship with national area, to correct for differences in definitions, to consult statistical agencies on likely uncertainties involved and to compare with international data sets. For the specific case of Colombia, the production of the oficial mangrove area statistics for the whole country must be lead for an unique institution; information used to get it must be derived from the same methodological approach, avoiding the current situation of different sources of information with unknowing uncertainties, different scales of analysis and use of remote sensing information from different years.

%
%\Jhoa{could be interesting to mention that IDEAM published a protocol for generating information on the distribution, extension and changes in forest cover of Colombia, from the Landsat image processing?}
%
Regarding to uncertainties related with the AGB estimation, homologation of approaches used for both field sampling and remote sensing tools are necessary. The implementation of a national mangrove inventory is probably the best way to ensure the use of the same approach and in this way reduce the current uncertainties. Nevertheless, in the case that mangrove national inventories are not contemplate, a general protocol to estimate biomass/carbon in mangroves should be proposed to ensure the use of a common methodological approach. The design of a national forest monitoring system for mangroves could also help to represent in a better way the variability of the carbon stocks in AGB due to differences between type of mangroves along both coast. %Currently, due to the study of  mangroves are linked to particular initiatives, there are locations in the country (i.e. Ci\'{e}naga Grande de Santa Marta,  Cispat\'{a} bay and M\'{a}laga bay) better represented than other. 
%Furthermore, increasing the number of representative sample plots and measurements, further stratification of estimates on the basis of mangrove types are also important ways to reduce the uncertainties associated with the estimation of this parameter.
%
%At country level important advances in terms of AGB estimation propose have been achieved \citep{Phillips2014}. Nevertheless, the contribution of this study in terms of uncertainty estimation could be helpful in case that the country decide to improve the AGB estimation resolution from the current scale level (Holdridge life zone classification) to ecosystem level.

@@

%Reduction of current levels of uncertainty in total AGB carbon stocks, requires reduction of uncertainties in AGB and mangrove area estimation. Ways to reduce those uncertainties are directly related with homogeneity of approaches to get information related with these input data.
The reduction of uncertainty in the estimation of the total carbon in Colombia is a major technical challenge for the inclusion of mangrove ecosystem in national mitigation strategies. This reduction could be achieve focussing the efforts in reducing uncertainty associated to area estimation or AGB estimation. This study suggest the best way to achieve this reduction for every coast; nevertheless, selection of mangrove area or AGB to reduce uncertainty is a decision which implies another issues related with capacities, logistic and financial resources.

>>>>

\begin{abstract}
Decision-making process, as well as the implementation of future climate change mitigation strategies for mangrove ecosystems, need the availability of reliable base information and tools to estimate the carbon stocks at regional and national levels. We estimated carbon (C) stocks in aboveground biomass (AGB) for mangroves in the Caribbean and Pacific coasts of Colombia. Using available data on AGB density and mangrove area for the whole country (excluding islands) and each coast independently, we estimated a national carbon stock in AGB for Colombian mangroves as 14.95 $\pm$ 2.72 TgC (mean $\pm$ SE), with 2.20 $\pm$ 0.86 TgC in the Caribbean coast and 9.61 $\pm$ 2.78 TgC in the Pacific coast. Uncertainty for total carbon in AGB in Colombian mangroves, reported as SE/mean in percentage, was 18\% at the national level, 39\% in the Caribbean coast, and 29\% in the Pacific coast. This uncertainty was more influenced by uncertainties associated with the estimation of mangrove area for the Caribbean coast, while for the Pacific coast it was more influenced by the uncertainties associated with AGB density. This difference is the result of a contrasting availability of AGB density data for both coasts.  Comparison between observed AGB density data and predictions from large-scale models showed that these models underestimate AGB density for Colombian mangroves. We reparameterized these models with our data, but found poor goodness-of-fit statistics for these model structures. We propose therefore three new statistical models to predict AGB density in Colombian mangroves based on climatic and vegetation data. In all cases, the best models included the enhanced vegetation index (EVI) and the mean temperature of driest quarter (BIO9). 
\end{abstract}

@@

The number of sources used reflects the scarce availability of public information in Colombia regarding biomass studies in mangrove ecosystems. Only those studies published with open access, or referenced in the Information System for the Management of Mangroves in Colombia (SIGMA) were used. Additionally, despite that more biomass inventories have been conducted in the Colombian mangroves, only those provided by INVEMAR could be accessed. The differences in terms of scale of the two sources of information, as well as the impossibility to include the different mangrove types along the two coasts, result in limitations in terms of the scope of our results. Nevertheless, considering the goal of the study, this first approach is a relevant step to identify the necessities of sampling in order to reduce the current uncertainties associated to the carbon stock estimation at national and/or regional level.

@@

%\begin{figure}[htbp]
%   \centering
%   \includegraphics{Figures/map.pdf} % requires the graphicx package
%\caption{Spatial distribution of mangroves in Colombia (green dots) and location of study sites with AGB density information (red stars)%.}
%   \label{fig:map}
%\end{figure}
%
Furthermore, information about mangrove cover area comes from two different sources: global estimations \citep{FAO2007, Giri2011, Giri2011dataset, CONL:CONL12060} and national estimations compiled by the Colombian institute for ocean and coastal research \citep{INVEMAR2014} and derived from particular estimations made by the different environmental authorities with jurisdiction in mangrove areas along the country. Those mangrove areas were estimated under different methodological approaches.

@@

Total carbon in AGB showed larger values in Pacific mangroves (9.61 $\pm$ 2.78 TgC) (mean $\pm$ SE) than in Caribbean mangroves (2.20 $\pm$ 0.86 TgC). This result is highly influenced by mangrove area, because if the comparison is made in terms of the same area unit (hectare), values of carbon density between both coasts do not show significant differences (Figure 1 %\ref{fig:Csotck}).

%\begin{figure}
%\includegraphics[%scale=0.4
%width=\textwidth
%]{Figures/Cdensity_and_TotalCstock.pdf} 
%\caption{
Figure 1. Aboveground biomass (AGB) density (Mg C ha$^{-1}$) and total carbon stock in AGB for Colombian mangroves (Tg C). 
%}
%\label{fig:Csotck}
%\end{figure}

@@

To achieve a 10\% uncertainty in AGB density for the Pacific coast, it would be necessary to establish at least 47 new sampling units, and a 5\% uncertainty may not be possible to achieve with less than 200 sampling units (Figure 2%\ref{fig:samplingEffort}). For the Caribbean coast the situation is less critical; it is possible to achieve a 5\% uncertainty in AGB density by adding 73 additional sampling units. Combined for the entire country, 81 additional sampling units are required to achieve 5\% uncertainty in AGB density. In all cases, we assume that these additional sampling units must follow the principles of random sampling.

@@

%\begin{figure} 
%\includegraphics[%scale=0.4
%width=\textwidth
%]{Figures/AGBsamplingEffort.pdf} 
%\caption{
Figure 2. Number of necessary sampling units of AGB density required to reduce current uncertainties at 10\% and 5\% level for Caribbean, Pacific and total Colombian mangroves.%} \label{fig:samplingEffort}
%\end{figure}

@@

We tested the ability of the models proposed by \citet{Twilley1992} and \citet{CONL:CONL12060}  in predicting our AGB density data. For both models, we found important overestimations of AGB density for the majority of the observations, although in some cases the models underestimated observed AGB density, particularly for values larger than 150-200 Mg ha$^{-1}$  (Figure 3 %\ref{fig:globalModels}).

@@

%\begin{figure} 
%\includegraphics[%scale=0.4
%width=\textwidth
%]{Figures/Globalmodelsvs.pdf} 
%\caption{
Fiigure 3. Observed vs. predicted AGB density (Mg ha$^{-1}$). Predicted values: using global models \citep{Twilley1992, CONL:CONL12060} and observed data: current data available for mangrove areas in Colombia}% \label{fig:globalModels}
%\end{figure}

@@

%\begin{figure} 
%\includegraphics[%scale=0.4
%width=\textwidth
%]{Figures/AGBdensity_map_final.pdf} 
%\caption{
Figure 4. Map of predicted AGB density (Mg ha$^{-1}$) based on the candidate models classified as ``substantial weight models''%} %\label{fig:Predicted AGB density map}
%\end{figure}

@@

%In general terms, to reduce uncertainties related with total AGB carbon stock estimation at coast level as well as at national level we propose the following ways in terms of area and AGB efforts, according with \citet{IPCC2003} recommendations:
%
Regarding the area estimation, the use the remote sensing data together with the ground based surveys approach  \citep{IPCC2003} has the advantage to be the most cost cost-efficient method of area measurement.Furthermore, in order to reduce the current uncertainties, it is important to check for a consistent relationship with the national area, to standardize definitions, to consult statistical agencies on likely uncertainties involved and to compare the results with international data sets. For the specific case of Colombia, the production of the official mangrove area statistics for the whole country must be led by an unique institution; the information used should derive from the same methodological approach, avoiding the current situation where different sources of information with unknowing uncertainties, different scales of analysis and use of remote sensing information from different years are used.

%
%\Jhoa{could be interesting to mention that IDEAM published a protocol for generating information on the distribution, extension and changes in forest cover of Colombia, from the Landsat image processing?}
%
Regarding the uncertainties related with the AGB estimation, homologation of approaches used for both field sampling and remote sensing tools are necessary. The implementation of a national mangrove inventory is probably the best way to ensure the use of the same approach and in this way reduce the current uncertainties. Nevertheless, in the case that mangrove national inventories are not contemplate, a general protocol to estimate biomass/carbon in mangroves should be proposed to ensure the use of a common methodological approach. The design of a national forest monitoring system for mangroves could also help to represent in a better way the variability of the carbon stocks in AGB due to differences between type of mangroves along both coasts. %Currently, due to the study of  mangroves are linked to particular initiatives, there are locations in the country (i.e. Ci\'{e}naga Grande de Santa Marta,  Cispat\'{a} bay and M\'{a}laga bay) better represented than other. 
%Furthermore, increasing the number of representative sample plots and measurements, further stratification of estimates on the basis of mangrove types are also important ways to reduce the uncertainties associated with the estimation of this parameter.
%
%At country level important advances in terms of AGB estimation propose have been achieved \citep{Phillips2014}. Nevertheless, the contribution of this study in terms of uncertainty estimation could be helpful in case that the country decide to improve the AGB estimation resolution from the current scale level (Holdridge life zone classification) to ecosystem level.

@@

%Reduction of current levels of uncertainty in total AGB carbon stocks, requires reduction of uncertainties in AGB and mangrove area estimation. Ways to reduce those uncertainties are directly related with homogeneity of approaches to get information related with these input data.
The reduction of uncertainty in the estimation of the total carbon in Colombia is a major technical challenge for the inclusion of mangrove ecosystem in national mitigation strategies. This reduction could be achieved focusing the efforts in reducing uncertainty associated to area estimation or AGB estimation. This study suggest the best way to achieve this reduction for each coast; nevertheless, selection of mangrove area or AGB to reduce uncertainty is a decision which implies other aspects related with capacities, logistic and financial resources.
